---
title: "**P8130 Final Report (Project 1)**"
author: 
- "Huanyu Chen (hc3451)   Xiaoting Tang (xt2288)"
- "Yifei Liu (yl5508)     Longyu Zhang (lz2951)"
output:
  pdf_document:
    latex_engine: pdflatex
fontsize: 11pt
linestretch: 1.5
---

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(MASS)
library(glmnet)
library(leaps)

set.seed(1)
```

# Abstract

(condenses a brief introduction, brief description of methods, and main
results into a one-paragraph summary)

# Introduction

The objective of this study is to use regression models to predict
academic performance in math, reading, and writing based on various
variables, including personal characteristics such as gender, ethnicity,
and parental education, as well as environmental factors like lunch
type, test preparation, and weekly study hours. Furthermore, the study
aims to identify potential correlations and regression model between
scores in different subjects. The combination of these analyses is
intended to provide educators and policymakers with practical insights
for tailoring interventions, improving educational programs, and
building strong support structures that promote students' overall
academic progress.

# Methods

This dataset provides information on public school students, including
three test scores and various personal and socioeconomic factors. To
facilitate analysis, categorical data have been converted to numerical
representations based on their ordinal order or type. We have excluded
the missing cells because they are factorial data types.

After processing the data, we created **Table 1**, which presents a
summary of the factorial data, including the number of missing data, the
number of categories under each variable, and the top counts. For the
numeric data (three test scores), we constructed a comprehensive
descriptive table (**Table 2**) to provide a snapshot of central
tendencies and variability. The distribution of the three response
variables (test scores) is presented in **Figure 1** (histogram) and
**Figure 2** (boxplot), indicating a normal distribution.

# Results

**Table 3**, **Table 4**, and **Table 5** display the regression models
for math, reading, and writing scores using both forward and backward
stepwise regression. Moreover, **Figures 3**, **Figures 4**, and
**Figures 5** display the diagnostic plots generated by the model.

# Conclusions/Discussion

# Contribution

**Xiaoting Tang**: Method, **Yifei Liu**: Result Display

**Longyu Zhang**: Interpretation, **Huanyu Chen**: Writing

# Appendix

## Table

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Read and Clean Data
data =
  read_csv("./data.csv") |>
  janitor::clean_names() |>
  mutate(
    gender = factor(case_when(
      gender == "male" ~ 0,
      gender == "female" ~ 1,
      )),
    ethnic_group = factor(case_when(
      ethnic_group == "group A" ~ 0,
      ethnic_group == "group B" ~ 1,
      ethnic_group == "group C" ~ 2,
      ethnic_group == "group D" ~ 3,
      ethnic_group == "group E" ~ 4,
      )),
    parent_educ = factor(case_when(
      parent_educ == "some highschool" ~ 0,
      parent_educ == "some college" ~ 1,
      parent_educ == "associate's degree" ~ 2,
      parent_educ == "bachelor's degree" ~ 3,
      parent_educ == "master's degree" ~ 4,
      )),
    lunch_type = factor(case_when(
      lunch_type == "standard" ~ 0,
      lunch_type == "free/reduced" ~ 1,
      )),
    test_prep = factor(case_when(
      test_prep == "none" ~ 0,
      test_prep == "completed" ~ 1,
      )),
    parent_marital_status = factor(case_when(
      parent_marital_status == "married" ~ 0,
      parent_marital_status == "single" ~ 1,
      parent_marital_status == "widowed" ~ 2,
      parent_marital_status == "divorced" ~ 3,
      )),
    practice_sport = factor(case_when(
      practice_sport == "never" ~ 0,
      practice_sport == "sometimes" ~ 1,
      practice_sport == "regularly" ~ 2,
      )),
    is_first_child = factor(case_when(
      is_first_child == "no" ~ 0,
      is_first_child == "yes" ~ 1,
      )),
    transport_means = factor(case_when(
      transport_means == "school_bus" ~ 0,
      transport_means == "private" ~ 1,
      )),
    wkly_study_hours = factor(case_when(
      wkly_study_hours == "< 5" ~ 0,
      wkly_study_hours == "10-May" ~ 1,
      wkly_study_hours == "> 10" ~ 2,
      ))
    ) |>
  mutate(nr_siblings = factor(nr_siblings))

# Another data set for EDA
data_long <- data |>
  pivot_longer(cols = c(math_score, reading_score, writing_score),
               names_to = "test", values_to = "score")
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Summary
sum_data_fct =
  data |>
  dplyr::select(1:11) |>
  skimr::skim() |>
  dplyr::select(skim_variable, n_missing, factor.n_unique, factor.top_counts)

colnames(sum_data_fct) = c("Variable", "Missing", "Unique", "Top Counts")

knitr::kable(x = sum_data_fct, caption = "Categorical Variables pre-analysis", digits = 1)

data =
  data |>
  drop_na()
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
sum_data_score =
  data |>
  dplyr::select(12:14) |>
  skimr::skim() |>
  dplyr::select(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100)

colnames(sum_data_score) = c("Variable", "Mean", "SD", "Min", "Q1", "Median", "Q3", "Max")

knitr::kable(x = sum_data_score, caption = "Continuous Variables pre-analysis", digits = 1)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# FULL MODEL
# Math
model_math_full = lm(math_score ~ . - reading_score - writing_score, data = data)

# Reading
model_reading_full = lm(reading_score ~ . - math_score - writing_score, data = data)

# Writing
model_writing_full = lm(writing_score ~ . - reading_score - math_score, data = data)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Stepwise Regressions
# math
math_sr = step(model_math_full, direction = 'both', trace = FALSE)

tb_math_sr = math_sr |>
  broom::tidy() |>
  filter(term != "(Intercept)") |> 
  dplyr::select(term, estimate, p.value)
colnames(tb_math_sr) = c("Term", "Estimate", "P Value")
knitr::kable(x = tb_math_sr, caption = "Math Scores Models by Stepwise Regression", digits = 2)

# reading
rea_sr = step(model_reading_full, direction = 'both', trace = FALSE)

tb_rea_sr = rea_sr |>
  broom::tidy() |>
  filter(term != "(Intercept)") |> 
  dplyr::select(term, estimate, p.value)
colnames(tb_rea_sr) = c("Term", "Estimate", "P Value")
knitr::kable(x = tb_rea_sr, caption = "Reading Scores Models by Stepwise Regression", digits = 2)

# writing
wri_sr = step(model_writing_full, direction = 'both', trace = FALSE)

tb_wri_sr = wri_sr |>
  broom::tidy() |>
  filter(term != "(Intercept)") |> 
  dplyr::select(term, estimate, p.value)
colnames(tb_wri_sr) = c("Term", "Estimate", "P Value")
knitr::kable(x = tb_wri_sr, caption = "Writing Scores Models by Stepwise Regression", digits = 2)
```

## Figure

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Histograms
data_long |>
  ggplot(aes(x = score, fill = test)) +
  geom_histogram(binwidth = 8, color = "#013571") +
  labs(
    title = "Figure 1: Scores Histograms by Subjects",
    x = "Score",
    y = "Count"
    ) +
  scale_fill_manual(values = c("#2E4E7D", "#405165", "#67A9CB")) +
  facet_grid(~ test) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Boxplots
data_long |>
  ggplot(aes(x = test, y = score, fill = test)) +
  geom_boxplot() +
  labs(
    title = "Figure 2: Scores Boxplot by Subjects",
    x = "Test",
    y = "Score"
    ) +
  facet_wrap(~ test, scales = "free") +
  scale_fill_manual(values = c("#2E4E7D", "#405165", "#67A9CB")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow = c(2, 2))
plot(math_sr, which = 1)
plot(math_sr, which = 2)
plot(math_sr, which = 3)
plot(math_sr, which = 4)
main_title <- "Figure 3: Diagnostic Plots for Math Test Score Regression Model"
mtext(main_title, side = 3, line = -2, outer = TRUE)


par(mfrow = c(2, 2))
plot(rea_sr, which = 1)
plot(rea_sr, which = 2)
plot(rea_sr, which = 3)
plot(rea_sr, which = 4)
main_title <- "Figure 4: Diagnostic Plots for Reading Test Score Regression Model"
mtext(main_title, side = 3, line = -2, outer = TRUE)


par(mfrow = c(2, 2))
plot(wri_sr, which = 1)
plot(wri_sr, which = 2)
plot(wri_sr, which = 3)
plot(wri_sr, which = 4)
main_title <- "Figure 5: Diagnostic Plots for Writing Test Score Regression Model"
mtext(main_title, side = 3, line = -2, outer = TRUE)
```
